---
title: "CYO: Video Game Sales with Ratings"
author: "Mohammad Awwad"
date: "1/19/2022"
output: pdf_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.align="center") 
```

```{r download knitr & rmarkdown, include=FALSE}
library(knitr)
library(rmarkdown)

```


<style>
body {
text-align: justify}
</style>

<br/><br/>

## Abstract

In this report, a video game sales in North America predictive model will be built based on selected training sets of the Video Game Sales with Ratings.  Video game sales from Vgchartz and corresponding ratings from Metacritic dataset are extracted from Kaggle which is an online community of data scientists and machine learners, owned by Google LLC.  Four algorithms will be applied: *Linear Regression*, *Polynomial Regression*, *Elastic Net* and *Random Forest*.  The result will be compared and analysed by the performance of Residual Mean Squared Error (RMSE).  

<br/><br/>

## 1. Introduction

Video game is an electronic game that involves interaction with a user interface to generate visual feedback on devices.  Playing video games is a kind of popular entertainment for both kids and adults.  The market is growing.   Publishers would like to predict video game sales for production and better allocation of limited resource.  Predictive model is to predict the video games sales in North America based on the Metascore in Metacritic which is a website that aggregates reviews of media products: films, TV shows, music albums, video games, and formerly, books.  Metascore is a weighted average of the most respected critics writing reviews online and in print.  The scores range from 0 to 100.  Scores below 20 represents overwhelming dislike, whereas scores over 90 represents universal acclaim.

Video Game Sales with Ratings dataset from Kaggle website (https://www.kaggle.com/rush4ratio/video-game-sales-with-ratings) will be used.  In this dataset, 80% of data is set as training data to build the predictive model and the other 20% of data is to evaluate the model by measuring Residual Mean Squared Error (RMSE).  Four algorithms are developed for comparison.

The goal of this project is to develop a machine learning algorithm to predict video game sales in North America based on the Metascore.  The lower the RMSE, the better the performance of the algorithm.

<br/><br/>
<P style="page-break-before: always">

## 2. Method

### 2.1 Data Cleaning

The source data was uploaded to Kaggle on Nov 2016.  

```{r Libray & File Download, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Install and Load Packages
if(!require(plotly)) install.packages("plotly", repos = "http://cran.us.r-project.org")
if(!require(corrplot)) install.packages("corrplot", repos = "http://cran.us.r-project.org")
if(!require(RCurl)) install.packages("RCurl", repos = "http://cran.us.r-project.org")
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(randomForest)) install.packages("randomForest", repos = "http://cran.us.r-project.org")
if(!require(kableExtra)) install.packages("kableExtra", repos = "http://cran.us.r-project.org")
library(dplyr)
library(ggplot2)
library(caret)
library(tidyr)
library(plotly)
library(RCurl)
library(corrplot)
library(randomForest)
library(kableExtra)
# Video Game Sales with Ratings
# Source File: https://www.kaggle.com/rush4ratio/video-game-sales-with-ratings
# File Download Path: https://github.com/mhmd-awwad/CYO/raw/main/Video_Games_Sales_as_at_22_Dec_2016.csv"
URL <- tempfile()
download.file("https://github.com/mhmd-awwad/CYO/raw/main/Video_Games_Sales_as_at_22_Dec_2016.csv",URL)
rawdata <- read.csv(file=URL)
```

The data type and summary statistics of each column of raw data downloaded are as follows:

```{r Raw Data Checking, echo=TRUE, message=FALSE, warning=FALSE, , echo=TRUE}
# Raw Data Checking: Type of each Column
str(rawdata)
# Raw Data Checking: Statistic of each column
summary(rawdata)
```

As the dataset were extracted on Nov 2016, records with "Year_of_Release" after 2016 are invalid.  Those records marked "NA" are also invalid.  These invalid records are required to be removed from the dataset.

```{r Data Cleansing-Year_of_Release, echo=TRUE, message=FALSE, warning=FALSE}
# Data Cleansing: Remove invalid records of "Year of Release" marked "NA"
cleandata <- rawdata %>% filter(!is.na(rawdata$Year_of_Release))
# Data generated in Nov 2016
# Data Cleansing: Change "Year of Release" to numeric and Remove invalid records of "Year of Release" after 2016
cleandata <- cleandata%>% dplyr::filter((as.numeric(as.character(cleandata$Year_of_Release)))<=2016)
```

The column "Rating" refers to the ESRB ratings.  No "RP" in this rating system and is to be replaced with correct rating.

```{r Data Cleansing: Correct record with wrong "Rating", echo=TRUE, message=FALSE, warning=FALSE}
# Data Cleansing: Correct record with wrong "Rating"
cleandata_rp <- cleandata %>% filter(Rating=="RP")
cleandata_rp
cleandata$Rating[cleandata$Rating == 'RP'] <- "E10+"
```

Those records with blank or "NA" rows are also removed from the dataset.

```{r Data Cleansing Invalid Records Removal, echo=TRUE, message=FALSE, warning=FALSE, , echo=TRUE}
# Data Cleansing: Remove invalid records of game with blank in "name"
cleandata <- cleandata %>% filter(cleandata$Name!="")
# Data Cleansing: Change "User_Score" to numeric
cleandata$User_Score <- as.numeric(as.character(cleandata$User_Score))
# Data Cleansing: Remove NA rows
finaldata <- na.omit(cleandata)
```
<P style="page-break-before: always">

### 2.2 Data Exploration

The structure of final dataset is as follows:

####2.2.1 No. of Records and no. of video games
```{r Data Exploratoin-No. of records, echo=TRUE, message=FALSE, warning=FALSE}
# Data Exploration: No. of rows and columns final dataset
dim(finaldata)
finaldata_record<-nrow(finaldata)
# Data Exploration: Statistic of each colum of final dataset
summary(finaldata)
# Data Exploration: No. of video games in final dataset
n_distinct(finaldata$Name)
game_no<-n_distinct(finaldata$Name)
```

The total number of records are **`r finaldata_record`**.

The total number of video games are **`r game_no`**.


####2.2.2 No. of Ratings by Genres

* Below chat shows top 5 genres are **Action**, **Shooter**, **Role-Playing**, **Sports** and **Racing**.  

```{r Data Exploration-Ratings by Genres, echo=TRUE, message=FALSE, warning=FALSE}
# Data Exploration: No. of Critic ratings in final dataset
finaldata_genres <- finaldata %>% group_by(Genre) %>%
  summarise(Critic_Rating=sum(Critic_Count)) %>%
  arrange(desc(Critic_Rating))
# Data Exploration: No. of Ratings by Genres Plot
finaldata_genres_p <-finaldata_genres%>%plot_ly(
  x = finaldata_genres$Genre,
  y = finaldata_genres$Critic_Rating,
  name = "Rating Distribution by Genres",
  type = "bar"
) %>% 
  add_text(text=finaldata_genres$Critic_Rating, hoverinfo='none', textposition = 'top', showlegend = FALSE,
           textfont=list(size=10, color="black"))%>%
  layout(xaxis = list(title = "Genres"),
         yaxis = list(title = "No. of Rating"))
finaldata_genres_p
```

####2.2.3 Top 10 video game with the greatest No. of Critic ratings (Metascore)

* Below chat shows top 5 video games with the greatest no. of Critic ratings (Metascore) are **Spider-Man 2**, **Grand Theft Auto V**, **Need for Speed: Most Wanted**, **Tomb Raider: Legend** and **Mass Effect 2**.  

```{r Data Exploration: Top 10 video game by Rating, echo=TRUE, message=FALSE, warning=FALSE}
# Data Exploration: Top 10 video game with the greatest No. of Critic ratings
finaldata_rating <- finaldata %>% group_by(Name) %>%
  summarize(Critic_Rating_Count=sum(Critic_Count)) %>%
  top_n(10) %>%
  arrange(desc(Critic_Rating_Count))
kable(finaldata_rating) %>%
  kable_styling(full_width = F) %>%
  column_spec(1, width = "20em")
```
<P style="page-break-before: always">

####2.2.4 Sales Trend in North America by Metascore

Below chart shows the sales volume in North America by Metascore

```{r Data Exploration: Sales in North America vs Critic Scores, echo=TRUE, message=FALSE, warning=FALSE}
# Data Exploration: Sales in North America vs Critic Scores
finaldata_NAsales <- finaldata %>%
  group_by(Critic_Score) %>%
  summarize(NA_Sales=sum(NA_Sales)) 
# Data Exploration: Sales in North America vs Critic Scores Plot
finaldata_NAsales_p <- plot_ly(finaldata_NAsales, x = ~finaldata_NAsales$Critic_Score, y = ~finaldata_NAsales$NA_Sales, type = 'scatter', mode = 'lines')%>%
  layout(xaxis = list(title = "Metascore"),
         yaxis = list(title = "Sales in North America (in millions of units)"))
finaldata_NAsales_p
```

In general, the sales volume of video game is higher with the higher the Metascore.


### 2.3 Create a train set and test set from final dataset

80% of final dataset will be set as training data and 20% of final dataset will be the testing data.

```{r Train set & Test set creation, echo=TRUE, message=FALSE, warning=FALSE, , echo=TRUE}
# test set will be 20% of finaldata
set.seed(1)
NASales_test_index <- createDataPartition(y = finaldata$NA_Sales, times = 1, p = 0.2, list = FALSE)
NASales_train_set <- finaldata[-NASales_test_index,]
NASales_test_set <- finaldata[NASales_test_index,]
```

### 2.4 RMSE Definitation 

Evaluation of prediction is based on Residual Mean Squared Error (RMSE).  RMSE is the typical error made when predicting sales in North America.  The lower the RMSE, the better the performance of the predication.   


```{r Define RMSE, echo=TRUE, message=FALSE, warning=FALSE} 
RMSE <- function(true_NA_Sales, predicted_NA_Sales){
  sqrt(mean((true_NA_Sales - predicted_NA_Sales)^2))
}
```


### 2.5 Models

#### 2.5.1  Model: Linear Regression
In this model, variable "Metascore" (Critic_Score) is to predict the sales in North America.  
```{r Model: Linear Regression, echo=TRUE, message=FALSE, warning=FALSE}
# Build the model on train dataset
lmModel <- lm(NA_Sales ~ Critic_Score, data=NASales_train_set) 
# Predict test dataset
lmPred <- predict(lmModel, NASales_test_set)  
# Model prediction performance
lm_rmse <- RMSE(lmPred, NASales_test_set$NA_Sales)
lm_rmse
# Create a Results Table
rmse_results <- data_frame(method = "Linear Regression", RMSE = lm_rmse)
rmse_results 
```

#### 2.5.2  Model: Polynomial Regression
A third-degree polynomial formula is developed in this model.
```{r Model: Polynomial Regression, echo=TRUE, message=FALSE, warning=FALSE}
# Build the model on train dataset
polyModel <- lm(NA_Sales ~ Critic_Score+ I(Critic_Score^2) + I(Critic_Score^3), data=NASales_train_set) 
# Predict test dataset
polyPred <- predict(polyModel, NASales_test_set)  
# Model prediction performance
poly_rmse <- RMSE(polyPred, NASales_test_set$NA_Sales)
poly_rmse
# Add Polynomial Regression result to the Results Table
rmse_results <- bind_rows(rmse_results,
                         data_frame(method="Polynomial Regression",  
                                    RMSE = poly_rmse))
```

#### 2.5.3  Elastic Net
Elastic Net is a penalized model which is effectively shrink coefficients and to set some coefficients to zero.
```{r Model: Elastic Net, echo=TRUE, message=FALSE, warning=FALSE}
# Build the model on train dataset
enModel <- train(
  NA_Sales~Critic_Score+ I(Critic_Score^2) + I(Critic_Score^3), data = NASales_train_set, method = "glmnet",
  trControl = trainControl("cv", number = 10),
  tuneLength = 10
)
# Model coefficients
coef(enModel$finalModel, enModel$bestTune$lambda)
# Make predictions
enPred<- enModel %>% predict(NASales_test_set)
# Model prediction performance
en_rmse <- RMSE(enPred, NASales_test_set$NA_Sales)
en_rmse
# Add Elastic net result to the Results Table
rmse_results <- bind_rows(rmse_results,
                         data_frame(method="Elastic Net",  
                                    RMSE = en_rmse))
```
<P style="page-break-before: always">

#### 2.5.4  Random Forest
Randon Forest is used to improve prediction performance and reduce instability by averaging multiple decision trees.
```{r Model: Random Forest, echo=TRUE, message=FALSE, warning=FALSE}
# Build the model on train dataset
rfModel <- randomForest(NA_Sales ~ Critic_Score, data = NASales_train_set, importance = TRUE)
# Predict test dataset
rfPred <- predict(rfModel, NASales_test_set)  
# Model prediction performance
rf_rmse <- RMSE(rfPred, NASales_test_set$NA_Sales)
rf_rmse
# Add Random Forest result to the Results Table
rmse_results <- bind_rows(rmse_results,
                         data_frame(method="Random Forest",  
                                    RMSE = rf_rmse))
```

<P style="page-break-before: always">

## 3. Results


###3.1 Result of Four Models

***3.1.1 Model: Linear Regression***

RMWE is *`r lm_rmse`*


***3.1.2 Model: Polynomial Regression***

RMSE is *`r poly_rmse`*


***3.1.3 Model: Elastic Net***

RMSE is *`r en_rmse`*


***3.1.4 Model: Random Forest***

RMSE is *`r rf_rmse`*

```{r RMSE Comparison Table, echo=TRUE, message=FALSE, warning=FALSE}
kable(rmse_results) %>%
  kable_styling(full_width = F) %>%
  column_spec(1, width = "20em")
```

The **best** model is **Random Forest** with **RMSE `r rf_rmse`**.

<br/><br/>

## 4. Conclusions
In this project, the Video Game Sales with Ratings dataset are used to build an algorithm to predict video game sales in North America based on the Metascore.  Four models, including "Linear Regression", "Polynomial Regression", "Elastic Net" and "Random Forest", are applied.  **"Random Forest"** got the best result, i.e. best RMSE, to predict video game sales in North America by Metascore.

